---
output: github_document
---

<!-- README.md is generated from README.Rmd. -->

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/wbelzak/regDIF.svg?branch=master)](https://travis-ci.org/wbelzak/regDIF)
[![Codecov test coverage](https://codecov.io/gh/wbelzak/regDIF/branch/master/graph/badge.svg)](https://codecov.io/gh/wbelzak/regDIF?branch=master)
<!-- badges: end -->

```{r, echo = FALSE}
options(width=80)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# regDIF: Regularized Differential Item Functioning

An R package that performs regularization of differential item functioning (DIF) parameters in item response theory (IRT) models using a penalized expectation-maximization algorithm.
 
## Version 1.0.0 Features

* Handles multiple continuous and categorical DIF covariates. 
* Supports binary and ordinal item responses. 
* Includes LASSO, ridge, MCP, and elastic net penalty functions for regularization.

## Installation

To get the current development version from Github:

```R
# install.packages("devtools")
devtools::install_github("wbelzak/regDIF")
```

## Getting Started

A simulated data example with 6 item responses (binary) and 3 background variables (gender, age, study) is used to demonstrate the `regDIF` package.

```{r, warning = FALSE, message = FALSE}
library(regDIF)
head(ida)
```

First, the item responses and predictor values are separately specified.

```{r, warning = FALSE, message = FALSE}
item.data <- ida[, 1:6]
pred.data <- ida[, 7:9]
```

Second, the `regDIF()` function fits a sequence of 10 tuning parameter values.

```{r, results = 'hide'}
fit <- regDIF(item.data, pred.data, num.tau = 10)
```

Finally, the DIF results are shown.

```{r}
summary(fit)
```

## Questions

Please send any questions to <wbelzak@gmail.com>.
