---
output: github_document
---

<!-- README.md is generated from README.Rmd. -->

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/wbelzak/regDIF.svg?branch=master)](https://travis-ci.org/wbelzak/regDIF)
<!-- badges: end -->

```{r, echo = FALSE}
options(width=80)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# regDIF: Regularized Differential Item Functioning

This R package performs regularization of differential item functioning (DIF) parameters in item response theory (IRT) models using a penalized expectation-maximization algorithm.
 
## Version 1.1.0 Features

regDIF can:

* Handle multiple continuous and categorical DIF covariates;
* Support binary, ordinal, and continuous item responses;
* Use LASSO, ridge, MCP, elastic net, and group penalty functions for regularization;
* Allow for proxy data to be used in place of estimating latent variable scores, which leads to
much faster estimation speed.

## Installation

To get the current released version from CRAN:

```R
install.packages("regDIF")
```

To get the current development version from Github:

```R
# install.packages("devtools")
devtools::install_github("wbelzak/regDIF")
```

## Getting Started

A simulated data example with 6 item responses (binary) and 3 background variables (gender, age, study) is available in the regDIF package:

```{r, warning = FALSE, message = FALSE}
library(regDIF)
head(ida)
```

First, the item responses and predictor values are separately specified:

```{r, warning = FALSE, message = FALSE}
item.data <- ida[, 1:6]
pred.data <- ida[, 7:9]
```

Second, the `regDIF()` function fits a sequence of 10 tuning parameter values using a penalized
EM algorithm, which assumes a normal latent variable affects all item responses:

```{r, results = 'hide'}
fit <- regDIF(item.data, pred.data, num.tau = 10)
```

The DIF results are shown below:

```{r}
summary(fit)
```

When estimation speed is slow, proxy data may be used in place of latent score estimation:
```{r, results = 'hide'}
fit_proxy <- regDIF(item.data, pred.data, prox.data = rowSums(item.data))
```
```{r}
summary(fit_proxy)
```

Other penalty functions (besides LASSO) may also be used. For instance, the elastic net penalty uses
a second tuning parameter, `alpha`, to vary the ratio of LASSO to ridge penalties:
```{r, results = 'hide'}
fit_proxy_net <- regDIF(item.data, pred.data, prox.data = rowSums(item.data), alpha = .5)
```
```{r}
summary(fit_proxy_net)
```

## Questions

Please send any questions to <wbelzak@gmail.com>.
